---
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# clust431

<!-- badges: start -->

<!-- badges: end -->

The goal of clust431 is to provide two main functions: k_means(), which
performs clustering on a given dataset using a basic k means algorithm
hclust(), which ...

## Installation

You can install the released version of clust431 from GitHub with:

``` r

install_github("josh-rowe/lab-7-clustering-methods-josh-rowe")
```

## Functions

### k_means()

#### Inputs

*dat* a dataframe where the first two columns are x and y coordinates

*k* a number describing the number of clusters to search for

*pca* a boolean which allows the user to have k_means() perform pca on
the data before running the k means algorithm (defaults to FALSE)

*max_iter* a number describing the maximum number of iterations for
k_means() to run through if convergence is not met (defaults to 100)

#### Outputs

A list with two elements:

*groups* a dataframe with the x and y values of the original data and a
factor which contains the grouping assignment of each observation

*tss* the total sum of squares of the final grouping assignment

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(clust431)

dat <- data.frame(x = c(1, 2, 3, 10, 11, 12, 20, 21, 22),
                  y = c(1, 2, 3, 10, 11, 12, 20, 21, 22))

dat_groups <- k_means(dat = dat, k = 3)


```

### hier_clust()

#### Inputs

*data* a numeric data frame

*cluster* a number describing the number of clusters to search for

#### Outputs

A data frame with two columns and a number of rows corresponding to the
inputted data frame:

*index* index row number corresponding to the inputted data set

*clusters* cluster assignment for index

## hier_clust() Example

This is a basic example which shows you how to solve a common problem:

```{r h_example}
library(clust431)

data <- iris %>% dplyr::select(-Species)
cluster <- 3

hierarchical <- hier_clust(data, cluster)
tail(hierarchical)
table(hierarchical$clusters)
```


